---
title: "NHANES Challenge"
format: html
editor: visual
---

## Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(haven)
library(here)
library(ggplot2)
library(janitor)
```

## Read-in Datasets for this Analysis:

-   DR1TOT_J.xpt as 'diet_behavior'
-   2017-2018_Hg-Blood.XPT as 'blood_hg'
-   2017-2018_Hg-Urine.XPT as 'urine_hg'
-   2017-2018_Diabetes.XPT as 'diabetes'

```{r}
diet_behavior <- read_xpt(here("data/nhanes_data/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes_data/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes_data/2017-2018_Demographics.XPT"))
```

## Subset Read-in Datasets

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

## Merge Subsets Into A Working Dataframe as 'df'

```{r}
df <- list(diet, tiid, bhg, uhg)

df <- df %>% reduce(full_join, by = 'SEQN')
```

Filter Dataframe df for the following:

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DIQ010 or DRD360 are NA
  filter(!is.na(DIQ010), !is.na(DRD370B)) %>%
  # Keep only rows where DIQ010 and DRD360 are 1 or 2
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DIQ010 = ifelse(DIQ010 == 1, "Has Diabetes", "No Diabetes"),
    DRD370B = ifelse(DRD370B == 1, "Consumes Ahi", "No Ahi")
  )
  
```

### Exploratory Data Analysis - On Your Own

-   Determine the prior probability that someone has diabetes (DIQ010)
    -   P(B) = 0.13

```{r}
#df %>% 
#  tabyl(DIQ010) %>% 
#  adorn_totals("row")
```

```{r}
#prior_probability_has_diabetes <- 0.13
```

-   Determine the prior probability that someone consumes ahi (DRD370B)
    -   P(A) = 0.36

```{r}
#df %>% 
#  tabyl(DRD370B) %>% 
#  adorn_totals("row")
```

```{r}
#prior_probability_consumes_ahi <- 0.36
```

-   Determine the likelihood of consuming ahi and having diabetes
    -   L(B\|A) = 0.36

```{r}
#df %>% 
#  tabyl(DRD370B, DIQ010) %>% 
#  adorn_percentages("col")
```

```{r}
#likelihood_consumes_ahi_has_diabetes <- 0.36
```

-   Determine posterior probability via bayes theorem
    -   P(B\|A) = 0.13

```{r}
#(prior_probability_has_diabetes * likelihood_consumes_ahi_has_diabetes) / prior_probability_consumes_ahi
```

-   What other questions can we ask of this data?

    -   How would an individual's race impact the prior probability of having diabetes or consuming ahi?

    -   How would an individual's age impact the prior probability of having diabetes or consuming ahi?

    -   How would an individual's gender impact the prior probability of having diabetes or consuming ahi?

    -   How would these varying demographic factors impact the likelihood of having diabetes and consuming ahi?

    -   How would these varying demographic factors impact the posterior probability of having diabetes and consuming ahi?

### Bayes Theorem - In Class

$$
P(B \mid A) = \frac{P(B) L(B \mid A)}{P(A)} 
$$

B \<- Has Diabetes

A \<- Consumes Ahi

#### P(B)

The probability that a person has diabetes. P(B) = 0.13

```{r}
df %>% 
  tabyl(DIQ010)
```

```{r}
probability_has_diabetes <- 0.13
```

#### L(B\|A) = P(A\|B)

The likelihood that a person has diabetes given that they consume ahi is equal to the probability that a person consumes ahi given that they have diabetes. L(B\|A) = 0.36

```{r}
df %>% 
  tabyl(DRD370B, DIQ010) %>% 
  adorn_percentages("col")
```

```{r}
probability_consumes_ahi_given_has_diabetes <- 0.36
```

#### P(A)

The probability that a person consumes ahi. P(A) = 0.36

```{r}
df %>% 
  tabyl(DRD370B)
```

```{r}
probability_consumes_ahi <- 0.36
```

#### P(B\|A) 

Posterior Probability: The probability that a person has diabetes given that they consume ahi. P(B\|A) = 0.13 \*Not much of a difference

```{r}
(probability_has_diabetes * probability_consumes_ahi_given_has_diabetes) / probability_consumes_ahi
```

```{r}
posterior_probability_has_diabetes_given_consumes_ahi <- 0.13
```

### Bayes Theorem - Diabetes & Age- On Your Own

$$
P(B \mid A) = \frac{P(B) L(B \mid A)}{P(A)} 
$$

B \<- Has Diabetes

A \<- Over 40

```{r}
demographics <- demographics %>% 
  select(SEQN, RIDRETH1, RIAGENDR, RIDAGEYR)
```

```{r}
nhanes_data <- left_join(df, demographics, by = "SEQN")
```

```{r}
nhanes_data <- nhanes_data %>% 
  mutate(age_bracket = ifelse(RIDAGEYR > 40, "Over 40", "Under 40"))
```

```{r}
ggplot(nhanes_data, aes(x = RIDAGEYR, fill = DIQ010)) +
  geom_density(alpha = 0.5)
```

#### P(B)

The probability that a person has diabetes. P(B) = 0.13

```{r}
nhanes_data %>% 
  tabyl(DIQ010)
```

```{r}
probability_has_diabetes
```

#### L(B\|A) = P(A\|B)

The likelihood that a person has diabetes given that they are over 40 years old is equal to the probability that a person is over 40 years old given that they have diabetes. L(B\|A) = 0.96

```{r}
nhanes_data %>% 
  tabyl(DIQ010, age_bracket) %>% 
  adorn_percentages("row")
```

```{r}
probability_over_40_given_has_diabetes <- 0.96
```

#### P(A)

The probability that a person is over 40 years old. P(A) = 0.53

```{r}
nhanes_data %>% 
  tabyl(age_bracket)
```

```{r}
probability_over_40 <- 0.53
```

#### P(BIA)

Posterior Probability: The probability that a person has diabetes given that they are over 40 years old. P(BIA) = 0.24

```{r}
(probability_has_diabetes * probability_over_40_given_has_diabetes) / probability_over_40
```

```{r}
posterior_probability_has_diabetes_given_over_40 <- 0.24
```
