---
title: "NHANES Challenge"
format: html
editor: visual
---

### Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(haven)
library(here)
library(ggplot2)
library(janitor)
```

### Read-in Datasets for this Analysis:

-   DR1TOT_J.xpt as 'diet_behavior'
-   2017-2018_Hg-Blood.XPT as 'blood_hg'
-   2017-2018_Hg-Urine.XPT as 'urine_hg'
-   2017-2018_Diabetes.XPT as 'diabetes'

```{r}
diet_behavior <- read_xpt(here("data/nhanes_data/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes_data/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes_data/2017-2018_Demographics.XPT"))
```

### Subset Read-in Datasets

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

### Merge Subsets Into A Working Dataframe as 'df'

```{r}
df <- list(diet, tiid, bhg, uhg)

df <- df %>% reduce(full_join, by = 'SEQN')
```

Filter Dataframe df for the following:

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DIQ010 or DRD360 are NA
  filter(!is.na(DIQ010), !is.na(DRD370B)) %>%
  # Keep only rows where DIQ010 and DRD360 are 1 or 2
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DIQ010 = ifelse(DIQ010 == 1, "Yes", "No"),
    DRD370B = ifelse(DRD370B == 1, "Yes", "No")
  )
  
```

Do some exploratory data analysis

-   Determine the prior probability that someone has diabetes (DIQ010)
    -   P(B) = 0.13

```{r}
df %>% 
  tabyl(DIQ010) %>% 
  adorn_totals("row")
```

```{r}
prior_probability_has_diabetes <- 0.13
```

-   Determine the prior probability that someone eats tuna (DRD370B)
    -   P(A) = 0.36

```{r}
df %>% 
  tabyl(DRD370B) %>% 
  adorn_totals("row")
```

```{r}
prior_probability_eats_tuna <- 0.36
```

-   Determine the likelihood of eating tuna and having diabetes
    -   L(B\|A) = 0.36

```{r}
df %>% 
  tabyl(DRD370B, DIQ010) %>% 
  adorn_percentages("col")
```

```{r}
likelihood_eats_tuna_has_diabetes <- 0.36
```

-   Determine posterior probability via bayes theorem
    -   P(B\|A) = 0.13

```{r}
(prior_probability_has_diabetes * likelihood_eats_tuna_has_diabetes) / prior_probability_eats_tuna
```

-   What other questions can we ask of this data?

    -   How would an individual's race impact the prior probability of having diabetes or eating tuna?

    -   How would an individual's age impact the prior probability of having diabetes or eating tuna?

    -   How would an individual's gender impact the prior probability of having diabetes or eating tuna?

    -   How would these varying demographic factors impact the likelihood of having diabetes and eating tuna?

    -   How would these varying demographic factors impact the posterior probability of having diabetes and eating tuna?
